---
title: Sorvi-tutorial
author: Louhos-team
---

Finnish open government data toolkit for R
===========

This is an R package for Finnish open government data. New
contributions are [welcome!](http://louhos.github.com/contact.html).

This work is part of the [rOpenGov](http://ropengov.github.com)
project.


## Installation

General users (CRAN release version):

```{r install, eval=FALSE}
install.packages("sorvi")
library(sorvi)
```

Developers (Github development version):

```{r install2, eval=FALSE}
install.packages("devtools")
library(devtools)
install_github("sorvi", "ropengov")
library(sorvi)
```

Further installation and development instructions can be found at the
project [home page](http://ropengov.github.com/sorvi).


## Using the package

For further usage
examples, see [Louhos-blog](http://louhos.wordpress.com) and
[Datawiki](https://github.com/louhos/sorvi/wiki/Data).


### Personal identification number (HETU)

Extracting information from a Finnish personal identification number:

```{r hetu, message=FALSE}
library(sorvi)
hetu("111111-111C")
```

Validating Finnish personal identification number:

```{r hetu2, fig.message=FALSE}
valid.hetu("010101-0101") # TRUE/FALSE
```

### Postal codes

Get Finnish postal codes vs. municipalities table from Wikipedia

```{r postalcodes, message=FALSE}
postal.code.table <- GetPostalCodeInfo()
head(postal.code.table)
```

### IP Location

Get geographic coordinates for a given IP-address from
http://www.datasciencetoolkit.org//ip2coordinates/

```{r iplocation, message=FALSE}
ip_location("137.224.252.10")
```


### Municipality information

Finnish municipality information is available through Population
Registry (Vaestorekisterikeskus), Statistics Finland (Tilastokeskus)
and Land Survey Finland (Maanmittauslaitos). We provide separate
download routine for each data set. The row names are in harmonized
format and can be used to match data sets from different sources, as
different data sets may carry slightly different versions of certain
municipality names. Examples for each case:

Finnish municipality information from Land Survey Finland ([Maanmittauslaitos, MML](http://www.maanmittauslaitos.fi/aineistot-palvelut/latauspalvelut/avoimien-aineistojen-tiedostopalvelu)).

```{r municipalityMML, message=FALSE, warning=FALSE}
municipality.info.mml <- GetMunicipalityInfoMML()
municipality.info.mml[1:2,]
```

Get information of Finnish provinces from Statistics Finland ([Tilastokeskus](http://pxweb2.stat.fi/Database/Kuntien%20perustiedot/Kuntien%20perustiedot/Kuntaportaali.px))

```{r municipalityStatFi, message=FALSE, warning=FALSE}
municipality.info.statfi <- GetMunicipalityInfoStatFi()
municipality.info.statfi[1:2,]
```

List the province for each municipality in Finland:
```{r province2, message=FALSE, warning=FALSE}

# Specific municipalities
m2p <- FindProvince(c("Helsinki", "Tampere", "Turku"))
head(m2p)

# All municipalities
m2p <- FindProvince(municipality.info.statfi$Kunta)

# Speeding up with predefined municipality info table:
m2p <- FindProvince(c("Helsinki", "Tampere", "Turku"), municipality.info.mml)
head(m2p)
```

Convert municipality codes and names:
```{r province3, message=FALSE}
municipality_ids <- ConvertMunicipalityCodes()
head(municipality_ids)
```

Translate municipality names Finnish/English:

```{r translate, message=FALSE}
translations <- LoadData("translations")
head(translations)
```

### Retrieve population register data

Municipality-level population information from [Vaestorekisterikeskus](http://vrk.fi/default.aspx?docid=5127&site=3&id=0):

```{r popreg, message=FALSE}
df <- GetPopulationRegister()
head(df)
```

### Province information

Get information of Finnish provinces from Wikipedia:

```{r municipality, message=FALSE}
tab <- GetProvinceInfoWikipedia()
head(tab)
```

### Company subsidies from the Finnish government

Finnish broadcasting company YLE published a large data set on Finnish company subsidies ([(C) MOT 10.9.2012](http://ohjelmat.yle.fi/mot/10_9) over 15 years. See the site for more information; CC-BY-SA 3.0-license.

```{r yritystuet-1, warning=FALSE, message=FALSE}
tuet <- GetMOTYritystuet()
head(tuet)
```


### Visualization routines

Line fit with confidence smoothers:

```{r regressionline, message=FALSE}
library(sorvi);
data(iris);
p <- vwReg(Sepal.Length ~ Sepal.Width, iris)
print(p)
```

Plot matrix:

```{r , message=FALSE}
mat <- rbind(c(1,2,3), c(1, 3, 1), c(4,2,2));
pm <- PlotMatrix(mat, "twoway", midpoint = 2)

# Plotting the scale
# sc <- PlotScale(pm$colors, pm$breaks)
```


# Väestörekisteritietojen visualisointi

Tämä esimerkki hakee Suomen kuntarajat ja vaestorekisterin asukaslukutiedot kunnittain, ja laskee sekä visualisoi sukupuolten suhteellisen osuuden eri kunnissa Suomen kartalla. Lähde: [Louhos-blogi](https://louhos.wordpress.com/2011/10/09/r-paketti-avoimen-suomi-datan-louhintaan/).

```{r popreg-2, message=FALSE, warning=FALSE, fig.width=8, fig.height=8, eval=TRUE}
library(sorvi)

# hae suomen kartta ja kuntarajat gadm-muodossa
library(gisfi)
gadm <- GetGADM("FIN_adm", "Kunta")

# vaestorekisterin asukasluvut kunnittain
vrek <- GetPopulationRegister()

# Liita vaestorekisterin tiedot karttaobjektiin ja
# aseta nollaan asukasluku kunnissa joiden osalta se ei ole tiedossa
gadm$asukkaita <- log10(rowSums(vrek[gadm$Kunta, c("Miehet", "Naiset")]))
gadm$asukkaita[is.na(gadm$asukkaita)] <- 0
# Laske myos sukupuolten suhteellinen osuus
gadm$miehet.osuus <- vrek[gadm$Kunta, "Miehet"]/vrek[gadm$Kunta, "Yhteensa"]
gadm$naiset.osuus <- vrek[gadm$Kunta, "Naiset"]/vrek[gadm$Kunta, "Yhteensa"]
# Aseta arvoon 50% miesten/naisten osuus
# kunnissa joiden osalta vakiluku ei ole tiedossa
gadm$miehet.osuus[is.na(gadm$miehet.osuus)] <- 0.5
gadm$naiset.osuus[is.na(gadm$naiset.osuus)] <- 0.5

# paletin rajapisteet
varname <- "naiset.osuus"
interval <- max(abs(gadm[[varname]] - 0.5))
at <- seq(0.5 - interval, 0.5 + interval, length = 100)

# Piirra Suomen kartta varitettyna naisten suhteellisen osuuden nojalla
q <- PlotShape(gadm, varname, type = "twoway",
      at = at, main = "Naiset Suomen kunnissa")
```
